### 2021-02-23 abstract.tex
Typeset Vampire in small caps

This paper describes and evaluates a symbol precedence recommender,
a machine learning system that estimates the best possible precedence
based on observations of prover performance on a set of problems and random precedences.
Using the graph convolutional neural network technology,
the system does not presuppose the problems to be related or share a common signature. 
When coupled with the theorem prover Vampire and evaluated on the \acrshort{tptp} problem library,
the recommender is found to outperform a state-of-the-art heuristic by more than \SI{4}{\percent}.

>>>>

This paper describes and evaluates a symbol precedence recommender,
a machine learning system that estimates the best possible precedence
based on observations of prover performance on a set of problems and random precedences.
Using the graph convolutional neural network technology,
the system does not presuppose the problems to be related or share a common signature. 
When coupled with the theorem prover \Vampire{} and evaluated on the \acrshort{tptp} problem library,
the recommender is found to outperform a state-of-the-art heuristic by more than \SI{4}{\percent}.

### 2021-02-23 evaluation.tex
Add a todo
Typeset Vampire in small caps

The empirical evaluation was performed using a modified version of the \gls{atp} Vampire 4.3.0 \cite{10.1007/978-3-642-39799-8_1}.
The prover was used to generate the training data and to evaluate the trained precedence recommender.
To generate the training data,
Vampire was modified to output \gls{cnf} representations of the problems
and annotated problem signatures in a machine-readable format.
%in \gls{json} format and annotated problem signatures in \gls{csv} format.\todo{MS: Mentioning JSON or CSV seems to be too low level. ``programmer documatation'' material.}
For the evaluation of the precedences generated by the recommender,
Vampire was modified to allow the user to supply explicit predicate and function symbol precedences for the proof search
(normally, the user only picks a precedence generation heuristic).
\todo[inline]{FB: Add a link to the modified Vampire.}

@@

Vampire was run with a time limit of 10 seconds.
To increase the potential impact of predicate precedences,
we used a simple \gls{tkbo} \cite{Ludwig2007,Kovacs2011}
that compares atoms according to the predicate precedence first,
Using the ordinary \gls{kbo} as a tie-breaker.
Ordinary \gls{kbo} was used to compare terms.
\todo[inline,author=FB]{Describe the detailed configuration in an appendix and reference the appendix here.}

@@

The input problems are assumed to be specified in the \gls{cnf} or the \gls{fof} fragment
of the \acrshort{tptp} language \cite{Sutcliffe2017}.
% http://www.tptp.org/TPTP/SyntaxBNF.html
\Gls{fof} problems are first converted into equisatisfiable \gls{cnf} problems by Vampire.
New symbols may be introduced during preprocessing,
namely due to the Tseitin transformation and Skolemization.
We denote the preprocessed \gls{cnf} problem as $P = (\Sigma, \mathit{Cl})$,
where $\Sigma$ denotes the list of all predicate and function symbols of the problem,
and $\mathit{Cl}$ denotes a set of clauses over $\Sigma$ that constitute the problem.

@@

The examples are generated by sampling the \gls{fol} part of the problem library \gls{tptp} v7.4.0 \cite{10.1007/978-3-030-29436-6_29}.
A total of \num{17053} problems formulated in \gls{fof} and \gls{cnf} are sampled.
The problems formulated in \gls{fof} are converted to \gls{cnf} by Vampire.
% vampire --mode clausify
Let $\ProblemsTptp$ denote the set of all \num{17053} \gls{cnf} problems available for training and evaluation.

@@

The examples are generated by iterative sampling of $\ProblemsTptp$.
In each iteration, a problem $P \in \ProblemsTptp$ is chosen and Vampire is executed twice on $P$
with two (uniformly) random predicate precedences and a common random function precedence.
The ``background'' random function precedence serves as additional noise (in addition to the variability 
contained in \gls{tptp}) and makes sure that the predicate precedence recommender
will not be able to rely on any specificities that would come from fixing function precedences in the training data.

The two executions are compared in terms of performance:
predicate precedence $\PrecBetter$ is recognized as better than predicate precedence $\PrecWorse$,
denoted as $\Better{\PrecBetter}{\PrecWorse}{P}$,
if the proof search finishes successfully with $\PrecBetter$ ($\SolverRun{P}{\Prec} = \top$)
and if the number of iterations of the saturation loop with $\PrecBetter$ is smaller than with $\PrecWorse$.
\todo{FB: Use notation for saturation loop iterations if the notation was established earlier. 
If failure is denoted by $\infty$, we can simply use $<$.}
If one of the two precedences is recognized as better,
the example $(P, \PrecBetter, \PrecWorse)$ is produced,
where $\PrecBetter$ is the better precedence,
and $\PrecWorse$ is the other precedence.
Otherwise, for example, if Vampire fails on both precedences, we go back to sampling another problem.

@@

\subsubsection{Metrics}
To analyze the training process,
we tracked the dynamics of three metrics on $\ProblemsTrain$ and $\ProblemsVal$:
loss, accuracy and solver performance.
The solver performance was estimated once per 10 training epochs by running Vampire
on \num{1000} problems of the respective problem set
and counting the number of problems successfully solved within the time limit.
Since the accuracy is a proxy measure of solver performance and the loss is a proxy measure of the accuracy,
we are interested in generalization from loss to accuracy and from both loss and accuracy to solver performance.
Similarly, we are interested in generalization from training set to validation set.
\todo{MS: kdyz se to tady tak pekne pripravi, asi ctenar ceka, ze se to pozdeji 
nejak zuzitkuje. Pokud se rozhnodnem loss/accuracy/ATP zavislost dal nerozebirat,
asi je tu navic i tento odstavec, nebo aspon jeho cast.}

@@

A predicate precedence recommender was trained on approximately \num{500000} examples,
% Half of the 1M predicate examples was used for validation.
and a function precedence recommender was trained on approximately \num{400000} examples.
% Half of the 800k function examples was used for validation.
For each problem $P \in \ProblemsVal$,
a pair\todo{MS: ja vim, ze sis to "vyplejtval" v predchozi vete,
ale kde jinde nez tady rict explicitne "predicate precendence" a "function precedence"?
Z toho "pair" ot nemusi byt jasny...}
 of symbol precedences was generated by the respective trained recommender
and Vampire was run using these precedences
with a wall clock time limit of 10 seconds.
% Note: We have mentioned the time limit above and here we repeat it for readers who only read Results.
The results are averaged over 5 runs to reduce the effect of noise due to the wall clock time limit.
As a baseline, the performance of Vampire with the \texttt{frequency} precedence heuristic was evaluated
with the same time limit.
For comparison, the two trained recommenders are evaluated separately,
with the predicate precedence recommender using the \texttt{frequency} heuristic to generate the function precedences, and vice versa.

To generate a precedence for a problem,
\todo{MS: tenhle odstavec mozna prehodit s predchozim?}
the recommender first converts the problem to a machine-friendly \gls{cnf} format,
then converts the \gls{cnf} to a graph,
then predicts symbol costs using the \gls{gcn} model
and finally orders the symbols by their costs to produce the precedence.
To simplify the experiment, the time limit is only imposed on the Vampire run
and excludes the time taken by the recommender to generate the precedence.
\todo{MS: and if we don't, this would be the perfect place to summarize
how long the just described procedure took on an average problem,
or how it (roughly) scales with the graph/cnf/tptp\_file size.}
When run with 2 threads,
the preprocessing of a single problem
takes at most at most \num[round-mode=places,round-precision=2]{1.262744486} seconds
for \SI{80}{\percent} of the problems
according to data measured on a random sample of \num{1000} problems.
% Analysis: https://docs.google.com/spreadsheets/d/1GujYNEtETpC3jk4iyENLjptv8mDmI5f1ZEhbmwtqnPM/edit?usp=sharing
% VML-715
\todo{MS: 1) single-core bych chapal jako default. tj netreba zminovat?
2) kterych 1000? To opet drazdi k otazkam. tj to bych asi taky vynechal, ale...
3) to maximum prece zalezi na tom, kterych 1000 to bylo.}
\todo{MS: Idea: kdyz je to linearni: co cas potrebny na recommending
vztahnout k casu klauzifikace? Jako rict, ze to trva x-krat dyl nez jen klauzifikovat?}

>>>>

The empirical evaluation was performed using a modified version of the \gls{atp} \Vampire{} 4.3.0 \cite{10.1007/978-3-642-39799-8_1}.
The prover was used to generate the training data and to evaluate the trained precedence recommender.
To generate the training data,
\Vampire{} was modified to output \gls{cnf} representations of the problems
and annotated problem signatures in a machine-readable format.
%in \gls{json} format and annotated problem signatures in \gls{csv} format.\todo{MS: Mentioning JSON or CSV seems to be too low level. ``programmer documatation'' material.}
For the evaluation of the precedences generated by the recommender,
\Vampire{} was modified to allow the user to supply explicit predicate and function symbol precedences for the proof search
(normally, the user only picks a precedence generation heuristic).
\todo[inline]{FB: Add a link to the modified \Vampire{}.}

@@

\Vampire{} was run with a time limit of 10 seconds.
To increase the potential impact of predicate precedences,
we used a simple \gls{tkbo} \cite{Ludwig2007,Kovacs2011}
that compares atoms according to the predicate precedence first,
Using the ordinary \gls{kbo} as a tie-breaker.
Ordinary \gls{kbo} was used to compare terms.
\todo[inline,author=FB]{Describe the detailed configuration in an appendix and reference the appendix here.}

@@

The input problems are assumed to be specified in the \gls{cnf} or the \gls{fof} fragment
of the \acrshort{tptp} language \cite{Sutcliffe2017}.
% http://www.tptp.org/TPTP/SyntaxBNF.html
\Gls{fof} problems are first converted into equisatisfiable \gls{cnf} problems by \Vampire{}.
New symbols may be introduced during preprocessing,
namely due to the Tseitin transformation and Skolemization.
We denote the preprocessed \gls{cnf} problem as $P = (\Sigma, \mathit{Cl})$,
where $\Sigma$ denotes the list of all predicate and function symbols of the problem,
and $\mathit{Cl}$ denotes a set of clauses over $\Sigma$ that constitute the problem.

@@

The examples are generated by sampling the \gls{fol} part of the problem library \gls{tptp} v7.4.0 \cite{10.1007/978-3-030-29436-6_29}.
A total of \num{17053} problems formulated in \gls{fof} and \gls{cnf} are sampled.
The problems formulated in \gls{fof} are converted to \gls{cnf} by \Vampire{}.
% vampire --mode clausify
Let $\ProblemsTptp$ denote the set of all \num{17053} \gls{cnf} problems available for training and evaluation.

@@

The examples are generated by iterative sampling of $\ProblemsTptp$.
In each iteration, a problem $P \in \ProblemsTptp$ is chosen and \Vampire{} is executed twice on $P$
with two (uniformly) random predicate precedences and a common random function precedence.
The ``background'' random function precedence serves as additional noise (in addition to the variability 
contained in \gls{tptp}) and makes sure that the predicate precedence recommender
will not be able to rely on any specificities that would come from fixing function precedences in the training data.

The two executions are compared in terms of performance:
predicate precedence $\PrecBetter$ is recognized as better than predicate precedence $\PrecWorse$,
denoted as $\Better{\PrecBetter}{\PrecWorse}{P}$,
if the proof search finishes successfully with $\PrecBetter$ ($\SolverRun{P}{\Prec} = \top$)
and if the number of iterations of the saturation loop with $\PrecBetter$ is smaller than with $\PrecWorse$.
\todo{FB: Use notation for saturation loop iterations if the notation was established earlier. 
If failure is denoted by $\infty$, we can simply use $<$.}
If one of the two precedences is recognized as better,
the example $(P, \PrecBetter, \PrecWorse)$ is produced,
where $\PrecBetter$ is the better precedence,
and $\PrecWorse$ is the other precedence.
Otherwise, for example, if \Vampire{} fails on both precedences, we go back to sampling another problem.
\todo{FB: Discuss: How come we mix examples in which both precedences terminate with examples in which only one precedence terminates?}

@@

\subsubsection{Metrics}
To analyze the training process,
we tracked the dynamics of three metrics on $\ProblemsTrain$ and $\ProblemsVal$:
loss, accuracy and solver performance.
The solver performance was estimated once per 10 training epochs by running \Vampire{}
on \num{1000} problems of the respective problem set
and counting the number of problems successfully solved within the time limit.
Since the accuracy is a proxy measure of solver performance and the loss is a proxy measure of the accuracy,
we are interested in generalization from loss to accuracy and from both loss and accuracy to solver performance.
Similarly, we are interested in generalization from training set to validation set.
\todo{MS: kdyz se to tady tak pekne pripravi, asi ctenar ceka, ze se to pozdeji 
nejak zuzitkuje. Pokud se rozhnodnem loss/accuracy/ATP zavislost dal nerozebirat,
asi je tu navic i tento odstavec, nebo aspon jeho cast.}

@@

A predicate precedence recommender was trained on approximately \num{500000} examples,
% Half of the 1M predicate examples was used for validation.
and a function precedence recommender was trained on approximately \num{400000} examples.
% Half of the 800k function examples was used for validation.
For each problem $P \in \ProblemsVal$,
a pair\todo{MS: ja vim, ze sis to "vyplejtval" v predchozi vete,
ale kde jinde nez tady rict explicitne "predicate precendence" a "function precedence"?
Z toho "pair" ot nemusi byt jasny...}
 of symbol precedences was generated by the respective trained recommender
and \Vampire{} was run using these precedences
with a wall clock time limit of 10 seconds.
% Note: We have mentioned the time limit above and here we repeat it for readers who only read Results.
The results are averaged over 5 runs to reduce the effect of noise due to the wall clock time limit.
As a baseline, the performance of \Vampire{} with the \texttt{frequency} precedence heuristic was evaluated
with the same time limit.
For comparison, the two trained recommenders are evaluated separately,
with the predicate precedence recommender using the \texttt{frequency} heuristic to generate the function precedences, and vice versa.

To generate a precedence for a problem,
\todo{MS: tenhle odstavec mozna prehodit s predchozim?}
the recommender first converts the problem to a machine-friendly \gls{cnf} format,
then converts the \gls{cnf} to a graph,
then predicts symbol costs using the \gls{gcn} model
and finally orders the symbols by their costs to produce the precedence.
To simplify the experiment, the time limit is only imposed on the \Vampire{} run
and excludes the time taken by the recommender to generate the precedence.
\todo{MS: and if we don't, this would be the perfect place to summarize
how long the just described procedure took on an average problem,
or how it (roughly) scales with the graph/cnf/tptp\_file size.}
When run with 2 threads,
the preprocessing of a single problem
takes at most at most \num[round-mode=places,round-precision=2]{1.262744486} seconds
for \SI{80}{\percent} of the problems
according to data measured on a random sample of \num{1000} problems.
% Analysis: https://docs.google.com/spreadsheets/d/1GujYNEtETpC3jk4iyENLjptv8mDmI5f1ZEhbmwtqnPM/edit?usp=sharing
% VML-715
\todo{MS: 1) single-core bych chapal jako default. tj netreba zminovat?
2) kterych 1000? To opet drazdi k otazkam. tj to bych asi taky vynechal, ale...
3) to maximum prece zalezi na tom, kterych 1000 to bylo.}
\todo{MS: Idea: kdyz je to linearni: co cas potrebny na recommending
vztahnout k casu klauzifikace? Jako rict, ze to trva x-krat dyl nez jen klauzifikovat?}

### 2021-02-23 introduction.tex
Typeset Vampire in small caps

Modern saturation-based Automatic Theorem Provers (ATPs) such as E \cite{SCV:CADE-2019}, SPASS \cite{DBLP:conf/cade/WeidenbachDFKSW09} 
or {\sc Vampire}~\cite{DBLP:conf/cav/KovacsV13}
employ the superposition calculus \cite{DBLP:journals/logcom/BachmairG94,DBLP:books/el/RV01/NieuwenhuisR01} as their underlying inference system.
Integrating flavours of resolution \cite{DBLP:books/el/RV01/BachmairG01}, paramodulation \cite{Robinson1983}, and 
the unfailing completion \cite{Bachmair89completionwithout}, superposition is a powerful calculus with 
a native support for equational reasoning. The calculus is parametrized by a simplification ordering on terms % and literals,
and uses it to constrain the applicability of inferences, with a significant impact on performance.

Both of the two main classes of simplification orderings used in practice,
the \acrlong*{kbo} \cite{Knuth1983}
and the \acrlong*{lpo} \cite{Kamin1980},
are specified with the help of a 
\emph{symbol precedence}, an ordering on the signature symbols. %.\footnote{KBO is further parameterized by symbol weights.}
% but our reference implementation in Vampire~\cite{DBLP:conf/cav/KovacsV13} 
% uses for efficiency reasons only weights equal to one \cite{DBLP:conf/cade/KovacsMV11} and so we do not consider this parameter here.}
While the superposition calculus is refutationally complete for any simplification ordering \cite{DBLP:journals/logcom/BachmairG94},
the choice of the precedence has a significant impact on how long it takes to solve a given problem.

@@

Choosing the next inference to apply is arguably the most important decision point in the state-of-the-art
\todo{FB: Add "saturation-based"?}
\glspl{atp} for \gls{fol}.
Since the most popular \glspl{atp} use the superposition calculus with literal selection \cite{DBLP:journals/logcom/BachmairG94},
the \gls{sot}, such as the \gls{kbo} \cite{Knuth1983},
plays an important role in restricting the inferences and thus guiding the proof search.
The \gls{kbo} is typically specified by a symbol precedence,
that is a permutation of non-logical symbols of the input problem.
Each of the commonly used precedence heuristics orders the symbols
lexicographically according to a small number of simple
\todo{FB: Is using the "simple" terminology from PAAR paper better than trying to find a better term?}
symbol properties,
such as the number of occurrences in the input problem,
% vampire --symbol_precedence frequency
the arity of the symbol, or
% vampire --symbol_precedence arity
%occurrence in the conjecture,
% vampire --symbol_precedence_boost goal
%number of occurrences in unit clauses, or
% vampire --symbol_precedence_boost units
whether or not the symbol was introduced in the process of Tseitin transformation \cite{Tseitin1983} or Skolemization \cite{Harrison2009}.
\todo{FB: Is citing Harrison better than nothing?}
% vampire --introduced_symbol_precedence top
Given that each of the heuristics used in practice, up to our knowledge, fails on some problems,
\todo{Mention that we tried already to find a good simple heuristic by ML.}
the general task of
finding a symbol precedence that guides the proof search efficiently on an arbitrary problem
remains to be intriguing.
This paper explores the possibility of tackling this task by training a neural network
on the results of running the \gls{atp} Vampire \cite{10.1007/978-3-642-39799-8_1} with randomly sampled precedences.
\todo{Ensure that the first paragraph fits on the first page.}

@@

Modern \gls{fol} \glspl{atp}, such as Vampire, E or SPASS,
solve \gls{fol} problems by iteratively applying the inference rules of superposition calculus.
Literal selection and equation orienting are employed to limit the ... while preserving the refutation completeness.
Both literal selection and equation orienting are determined by \gls{sot}.

@@

The state-of-the-art \gls{fol} \glspl{atp},
such as \vampire{} \cite{10.1007/978-3-642-39799-8_1} and E \cite{10.1007/978-3-030-29436-6_29},
are saturation-based and
infer new provable clauses by applying the rules of the \definiendum{superposition calculus}.
\todo[author=Filip]{Cite some source that explains superposition calculus.}
The inferences under superposition calculus are constrained by a \definiendum{simplification ordering on the terms}.
\todo[author=Filip]{Do we mean terms in sensu stricto (applications of functions), or generalized terms (applications of functions and predicates)?}
The two classes of simplification orderings commonly used in practice are
\gls{lpo} \cite{Kamin1980} and \gls{kbo} \cite{Knuth1983}.
In both classes the ordering is specified by a \definiendum{symbol precedence},
\todo[author=Filip]{This may be an oversimplification: KBO is further parameterized by symbol weights.}
which is a permutation of predicate and function symbols of the input problem.

@@

The performance of the state-of-the-art \acrfullpl{atp} is severely limited by some of the design decisions that constitute the provers' architecture.
For example, the \frequency{} symbol precedence heuristic of the \acrfull{fol} \acrshort{atp} \vampire{} is easily outperformed by the best of 10 random precedences.\cite{}
The gap in performance suggests that a more elaborate precedence heuristic with a superior performance exists.
% what is precedence heuristic?
% Thesis statement
Training a model from proof attempts with random precedences yields a heuristic that significantly outperforms the state-of-the-art heuristic.

>>>>

Modern saturation-based Automatic Theorem Provers (ATPs) such as E \cite{SCV:CADE-2019}, SPASS \cite{DBLP:conf/cade/WeidenbachDFKSW09} 
or \Vampire{} \cite{DBLP:conf/cav/KovacsV13}
employ the superposition calculus \cite{DBLP:journals/logcom/BachmairG94,DBLP:books/el/RV01/NieuwenhuisR01} as their underlying inference system.
Integrating flavours of resolution \cite{DBLP:books/el/RV01/BachmairG01}, paramodulation \cite{Robinson1983}, and 
the unfailing completion \cite{Bachmair89completionwithout}, superposition is a powerful calculus with 
a native support for equational reasoning. The calculus is parametrized by a simplification ordering on terms % and literals,
and uses it to constrain the applicability of inferences, with a significant impact on performance.

Both of the two main classes of simplification orderings used in practice,
the \acrlong*{kbo} \cite{Knuth1983}
and the \acrlong*{lpo} \cite{Kamin1980},
are specified with the help of a 
\emph{symbol precedence}, an ordering on the signature symbols. %.\footnote{KBO is further parameterized by symbol weights.}
% but our reference implementation in \Vampire{}~\cite{DBLP:conf/cav/KovacsV13} 
% uses for efficiency reasons only weights equal to one \cite{DBLP:conf/cade/KovacsMV11} and so we do not consider this parameter here.}
While the superposition calculus is refutationally complete for any simplification ordering \cite{DBLP:journals/logcom/BachmairG94},
the choice of the precedence has a significant impact on how long it takes to solve a given problem.

@@

Choosing the next inference to apply is arguably the most important decision point in the state-of-the-art
\todo{FB: Add "saturation-based"?}
\glspl{atp} for \gls{fol}.
Since the most popular \glspl{atp} use the superposition calculus with literal selection \cite{DBLP:journals/logcom/BachmairG94},
the \gls{sot}, such as the \gls{kbo} \cite{Knuth1983},
plays an important role in restricting the inferences and thus guiding the proof search.
The \gls{kbo} is typically specified by a symbol precedence,
that is a permutation of non-logical symbols of the input problem.
Each of the commonly used precedence heuristics orders the symbols
lexicographically according to a small number of simple
\todo{FB: Is using the "simple" terminology from PAAR paper better than trying to find a better term?}
symbol properties,
such as the number of occurrences in the input problem,
% vampire --symbol_precedence frequency
the arity of the symbol, or
% vampire --symbol_precedence arity
%occurrence in the conjecture,
% vampire --symbol_precedence_boost goal
%number of occurrences in unit clauses, or
% vampire --symbol_precedence_boost units
whether or not the symbol was introduced in the process of Tseitin transformation \cite{Tseitin1983} or Skolemization \cite{Harrison2009}.
\todo{FB: Is citing Harrison better than nothing?}
% vampire --introduced_symbol_precedence top
Given that each of the heuristics used in practice, up to our knowledge, fails on some problems,
\todo{Mention that we tried already to find a good simple heuristic by ML.}
the general task of
finding a symbol precedence that guides the proof search efficiently on an arbitrary problem
remains to be intriguing.
This paper explores the possibility of tackling this task by training a neural network
on the results of running the \gls{atp} \Vampire{} \cite{10.1007/978-3-642-39799-8_1} with randomly sampled precedences.
\todo{Ensure that the first paragraph fits on the first page.}

@@

Modern \gls{fol} \glspl{atp}, such as \Vampire{}, E or SPASS,
solve \gls{fol} problems by iteratively applying the inference rules of superposition calculus.
Literal selection and equation orienting are employed to limit the ... while preserving the refutation completeness.
Both literal selection and equation orienting are determined by \gls{sot}.

@@

The state-of-the-art \gls{fol} \glspl{atp},
such as \Vampire{} \cite{10.1007/978-3-642-39799-8_1} and E \cite{10.1007/978-3-030-29436-6_29},
are saturation-based and
infer new provable clauses by applying the rules of the \definiendum{superposition calculus}.
\todo[author=Filip]{Cite some source that explains superposition calculus.}
The inferences under superposition calculus are constrained by a \definiendum{simplification ordering on the terms}.
\todo[author=Filip]{Do we mean terms in sensu stricto (applications of functions), or generalized terms (applications of functions and predicates)?}
The two classes of simplification orderings commonly used in practice are
\gls{lpo} \cite{Kamin1980} and \gls{kbo} \cite{Knuth1983}.
In both classes the ordering is specified by a \definiendum{symbol precedence},
\todo[author=Filip]{This may be an oversimplification: KBO is further parameterized by symbol weights.}
which is a permutation of predicate and function symbols of the input problem.

@@

The performance of the state-of-the-art \acrfullpl{atp} is severely limited by some of the design decisions that constitute the provers' architecture.
For example, the \frequency{} symbol precedence heuristic of the \acrfull{fol} \acrshort{atp} \Vampire{} is easily outperformed by the best of 10 random precedences.\cite{}
The gap in performance suggests that a more elaborate precedence heuristic with a superior performance exists.
% what is precedence heuristic?
% Thesis statement
Training a model from proof attempts with random precedences yields a heuristic that significantly outperforms the state-of-the-art heuristic.

### 2021-02-23 preliminaries.tex
Typeset Vampire in small caps

A \emph{\acrfull{fol} problem} consists of a set of axiom formulas and a conjecture formula.
In a \emph{refutation-based} \emph{\acrfull{atp}},
proving that the axioms entail the conjecture
is reduced to proving that the axioms together with the negated conjecture entail a \emph{contradiction}.
The most popular \gls{fol} \glspl{atp}, such as Vampire \cite{10.1007/978-3-642-39799-8_1}, E \cite{10.1007/978-3-030-29436-6_29}, or SPASS \cite{DBLP:conf/cade/WeidenbachDFKSW09},
start the proof search by converting the input \gls{fol} formulas to an equisatisfiable representation in 
% \emph{quantifier-free first-order 
\emph{\acrfull{cnf}} \cite{DBLP:books/el/RV01/NonnengartW01,Harrison2009}.
We denote a problem in \gls{cnf} as $P = (\Sigma, \mathit{Cl})$,
where $\Sigma$ is a list of all non-logical (predicate and function)
\emph{symbols} in the problem called the \emph{signature},
and $\mathit{Cl}$ is the set of the clauses of the problem (including the negated conjecture).

@@

influences the proof search in Vampire on two levels.
First, the inferences on each clause are limited
to the selected literals.
\todo{Cite Bachmair Ganzinger or Handbook of AR, chapter 2. For literal selection. Inspiration: Selecting the selection.}
In each clause,
either a negative literal or all the maximal literals are selected.
The maximality is evaluated
according to the simplification ordering.
% Note: The selection function Total does not use the simplification ordering.
Second, the simplification ordering orients some of the equalities
to prevent superposition and equality factoring
from inferring redundant complex conclusions.
In each of these two roles,
the simplification ordering may impact the direction and,
in effect, the length of the proof search.

>>>>

A \emph{\acrfull{fol} problem} consists of a set of axiom formulas and a conjecture formula.
In a \emph{refutation-based} \emph{\acrfull{atp}},
proving that the axioms entail the conjecture
is reduced to proving that the axioms together with the negated conjecture entail a \emph{contradiction}.
The most popular \gls{fol} \glspl{atp}, such as \Vampire{} \cite{10.1007/978-3-642-39799-8_1}, E \cite{10.1007/978-3-030-29436-6_29}, or SPASS \cite{DBLP:conf/cade/WeidenbachDFKSW09},
start the proof search by converting the input \gls{fol} formulas to an equisatisfiable representation in 
% \emph{quantifier-free first-order 
\emph{\acrfull{cnf}} \cite{DBLP:books/el/RV01/NonnengartW01,Harrison2009}.
We denote a problem in \gls{cnf} as $P = (\Sigma, \mathit{Cl})$,
where $\Sigma$ is a list of all non-logical (predicate and function)
\emph{symbols} in the problem called the \emph{signature},
and $\mathit{Cl}$ is the set of the clauses of the problem (including the negated conjecture).

@@

influences the proof search in \Vampire{} on two levels.
First, the inferences on each clause are limited
to the selected literals.
\todo{Cite Bachmair Ganzinger or Handbook of AR, chapter 2. For literal selection. Inspiration: Selecting the selection.}
In each clause,
either a negative literal or all the maximal literals are selected.
The maximality is evaluated
according to the simplification ordering.
% Note: The selection function Total does not use the simplification ordering.
Second, the simplification ordering orients some of the equalities
to prevent superposition and equality factoring
from inferring redundant complex conclusions.
In each of these two roles,
the simplification ordering may impact the direction and,
in effect, the length of the proof search.
